Solving Collective Spin Hamiltonians with Decoherence

### 2018-08-23 soc_squeezing/cumulant_expansions.tex
rearranged appendices and started generalizing the decoherence operator

Previous work has shown how to realize various spin-squeezing
Hamiltonians with ultracold atomic systems.  While it is
straightforward to simulate the dynamics generated by these
Hamiltonians in the context of a closed quantum system, here we try to
develop an understanding of their the squeezing behavior in the
presence of decoherence.  In particular, we will be considering
dynamics induced by the collective one-axis twisting (OAT), two-axis
twisting (TAT), and transverse field (TVF) Hamiltonians of the form
\begin{align}
  H_{\t{OAT}} = \chi S_z^2,
  &&
  H_{\t{TAT}} = \chi \p{S_z^2 - S_y^2},
  &&
  H_{\t{TVF}} = \chi S_z^2 - \Omega S_x,
\end{align}
for spins which decay in an uncorrelated manner at a rate $\gamma$.
The evolution of any Heisenburg operator $\O$ with a Hamiltonian $H$
is given by
\begin{align}
  \f{d}{dt} \O
  = i\sp{H,\O}_- + \gamma\D\p{\O},
  &&
  \D\p{\O} \equiv \sum_j\p{\sigma_j^+\O\sigma_j^-
    - \f12\sp{\sigma_j^+\sigma_j^-,\O}_+},
  \label{eq:EOM}
\end{align}
where $\sp{A,B}_\pm\equiv AB\pm BA$ and $j$ indexes an individual
spin.  Our measure of squeezing is
\begin{align}
  \xi^2 \equiv \f{N}{\abs{\bk{\v S}}^2}
  \min_{\uv n\perp\bk{\v S}} \bk{\p{\v S\c\uv n}^2},
  \label{eq:squeezing}
\end{align}
where $N$ is the total number of spins, $\v S$ is the collective spin
vector, and the minimization is performed over all unit vectors
$\uv n$ in the plane orthogonal to the mean spin vector $\bk{\v S}$.
This squeezing parameter is entirely determined by components
$\bk{S_\alpha}$ of the mean spin vector in addition to collective
spin-spin correlators of the form $\bk{S_\alpha S_\beta}$; our task is
therefore to compute these expectation values in all scenarios of
interest.

@@

The case of the one-axis twisting is simplest to consider, as we can
write the Hamiltonian in the form
\begin{align}
  H_{\t{OAT}} = \f14 \chi \sum_{j,k} \sigma_j^z \sigma_k^z
  = \f12 \chi \sum_{j<k} \sigma_j^z \sigma_k^z + \f14 N \chi,
\end{align}
which admits an exact solution previously worked out by Foss-Feig et
al.~in ref.~\cite{foss-feig2013nonequilibrium}.  Adapting exact
expectation values with respect to the initial state $\ket{X}$ which
has all spins pointing in $x$, in terms of $\mu,\nu\in\set{+1,-1}$ we
find that
\begin{align}
  \bk{S_+}
  &= \sum_j\bk{\sigma_j^+}
  = \f{N}{2} e^{-\gamma t/2} \Phi\p{\chi,t}^{N-1}, \label{eq:S+_OAT} \\
  \bk{S_\mu S_z}
  &= \f12\sum_j\bk{\sigma_j^\mu\sigma_j^z}
  + \f12\sum_{j\ne k} \bk{\sigma_j^\mu \sigma_k^z} \\
  &= -\f{\mu}{2}\bk{S_\mu} + \f14 N \p{N-1} e^{-\gamma t/2}
  \Psi\p{\mu\chi,t} \Phi\p{\chi,t}^{N-2}, \\
  \bk{S_\mu S_\nu}
  &= \sum_j \bk{\sigma_j^\mu \sigma_j^\nu}
  + \sum_{j\ne k} \bk{\sigma_j^\mu \sigma_k^\nu} \\
  &= \delta_{\mu,-\nu} \p{\f{N}{2} + \mu \bk{S_z}}
  + \f14 N \p{N-1} e^{-\gamma t}
  \Phi\p{\sp{\mu+\nu}\chi,t}^{N-2}, \label{eq:SS+-_OAT}
\end{align}
where
\begin{align}
  \Phi\p{\zeta,t} \equiv e^{-\gamma t/2}
  \sp{\cos\p{s_\zeta t}+\f12\gamma t~\t{sinc}\p{s_\zeta t}},
  &&
  \Psi\p{\zeta,t} \equiv e^{-\gamma t/2}
  \p{is_\zeta-\gamma/2}t~ \t{sinc}\p{s_\zeta t},
\end{align}
and $s_\zeta \equiv \zeta + i2\gamma$.  The last ingredients we need
to compute spin squeezing at any time are $\bk{S_z}$ and $\bk{S_z^2}$.
As these operators commute with the OAT Hamiltonian, their evolution
is governed entirely by decoherence, such that
\begin{align}
  \f{d}{dt}\p{S_z^n}
  = \gamma \D\p{S_z^n}
  = \gamma \p{S + S_z} \sum_{k=0}^{n-1}
  \p{-1}^{n-k} { n \choose k } S_z^k,
\end{align}
where $S\equiv N/2$ and $\D\p{S_z^n}$ is worked out in Appendix
\ref{sec:decoherence}.  In particular,
\begin{align}
  \f{d}{dt} S_z = -\gamma \p{S + S_z},
  &&
  \f{d}{dt}\p{S_z^2} = \gamma \p{S + S_z} \p{1 - 2S_z}.
\end{align}
The initial conditions $\bk{S_z}_{t=0}=0$ and $\bk{S_z^2}_{t=0}=S/2$
then imply
\begin{align}
  \bk{S_z} = S\p{e^{-\gamma t}-1},
  &&
  \bk{S_z^2} = S \p{1 - \f12 e^{-\gamma t}} e^{-\gamma t} + \bk{S_z}^2.
  \label{eq:Sz_OAT}
\end{align}
The expectation values in \eqref{eq:S+_OAT}-\eqref{eq:SS+-_OAT} and
\eqref{eq:Sz_OAT} are sufficient to compute the squeezing parameter
$\xi^2$ in \eqref{eq:squeezing} at any time throughout evolution of
the initial state $\ket{X}$ under $H_{\t{OAT}}$.

@@

The equations of motion for $S_\mu$, $S_\mu S_z$, and $S_\mu S_\nu$
(i.e.~where $\mu,\nu\in\set{+1,-1}$) under the OAT Hamiltonian
$H_{\t{OAT}}=\chi S_z^2$ are determined by
\begin{align}
  \f{d}{dt} \p{S_+^\ell S_-^m S_z^n}
  = i\chi \sp{S_z^2, S_+^\ell S_-^m S_z^n}_-
  + \gamma \D\p{S_+^\ell S_-^m S_z^n},
\end{align}
where we can use identity \eqref{eq:comm_z_mu_nu_pow} to simplify
\begin{align}
  \sp{S_z^2, S_+^\ell S_-^m S_z^n}_-
  = \sp{S_z^2, S_+^\ell S_-^m}_- S_z^n
  = S_+^\ell S_-^m \p{\sp{\ell-m}^2 + 2\sp{\ell-m} S_z} S_z^n,
\end{align}
and the effect of decoherence through $\D\p{S_+^\ell S_-^m S_z^n}$ is
worked out in Appendix \ref{sec:decoherence}.  In order to compute
squeezing of the OAT Hamiltonian in the presence of single-spin decay,
we therefore need to solve the systems of equations defined by
\begin{align}
  \f{d}{dt} \p{S_\mu S_z^n}
  = i\chi S_\mu \p{1 + 2\mu S_z} S_z^n
  + \gamma S_\mu \sp{\p{S - \delta_{\mu,-1} + S_z} \sum_{k=0}^{n-1}
    \p{-1}^{n-k} { n \choose k } S_z^k - \f12 S_z^n}
\end{align}
and
\begin{multline}
  \f{d}{dt} \p{S_\mu S_\nu S_z^n}
  = i\chi S_\mu S_\nu \p{\sp{\mu+\nu}^2 + 2 \sp{\mu+\nu} S_z} S_z^n \\
  + \gamma S_\mu S_\nu
  \sp{\p{S - \delta_{\mu,-1} - \delta_{\nu,-1} + S_z} \sum_{k=0}^{n-1}
    \p{-1}^{n-k} { n \choose k } S_z^k - S_z^n},
\end{multline}
for $\mu\ge\nu$.  The initial conditions for these systems of equations can be determined from the expansions
\begin{align}
  S_\mu = \sum_{m=-S}^S \sqrt{\p{S-\mu m}\p{S+\mu m+1}} \op{m+\mu}{m},
  &&
  S_z^n = \sum_{m=-S}^S m^n \op{m},
\end{align}
and the initial state
\begin{align}
  \ket{X} = 2^{-N/2} \sum_{m=-S}^S { N \choose S+m }^{1/2} \ket{m},
\end{align}
from which we can determine
\begin{align}
  \bk{S_\mu S_z^n}_{t=0}
  &= 2^{-N} \sum_{m=-S}^S m^n
  \sp{{ N \choose S+m } { N \choose S+m+\mu }
    \p{S-\mu m}\p{S+\mu m+1}}^{1/2} \\
  &= 2^{-N} \sum_{m=-S}^S m^n { N \choose S+m } \abs{m - \mu S},
\end{align}
and
\begin{multline}
  \bk{S_\mu S_\nu S_z^n}_{t=0}
  = 2^{-N} \sum_{m=-S}^S m^n
  \sp{{ N \choose S+m } { N \choose S+m+\mu+\nu }}^{1/2} \\
  \times \sp{\p{S-\nu\sp{m+\nu}}\p{S+\mu\sp{m+\nu}+1}
    \p{S-\nu m}\p{S+\nu m+1}}^{1/2}.
\end{multline}

\section{One-axis twisting with a transverse field}

We now consider again squeezing the initial state $\ket{X}$, but this
time with the transverse field Hamiltonian
\begin{align}
  H_{\t{TVF}}
  \equiv \chi S_z^2- \Omega S_x.
\end{align}
To determine the coherent evolution of the operator
$S_\mu^m S_\nu^n S_z^n$, we now need to compute
\begin{align}
  \sp{H_{\t{TVF}}, S_\mu^m S_\nu^n S_z^n}_-
  = \chi \sp{S_z^2, S_\mu^m S_\nu^n S_z^n}_-
  - \Omega \sp{S_x, S_\mu^m S_\nu^n S_z^n}_-.
\end{align}
Here first commutator is identical to that for OAT, while

@@

This text makes ubiquitous use of various spin operator identities; we
collect and derive particularly useful identities here for reference.
The elementary commutation relations between single-weight spin
operators are
\begin{align}
  \f12 \sp{\sigma_j^z,\sigma_k^\mu}_-
  &= \delta_{jk} \mu \sigma_j^\mu,
  &
  \sp{S_z,\sigma_j^\mu}_-
  &= \f12 \sp{\sigma_j^z,S_\mu}_- = \mu \sigma_j^\mu,
  &
  \sp{S_z,S_\mu}_-
  &= \mu S_\mu,
  \label{eq:comm_z_mu} \\
  \sp{\sigma_j^\mu,\sigma_k^\nu}_-
  &= \delta_{jk} \delta_{\mu,-\nu} \mu \sigma_j^z,
  &
  \sp{S_\mu,\sigma_j^\nu}_-
  &= \sp{\sigma_j^\mu,S_\nu}_-
  = \delta_{\mu,-\nu} \mu \sigma_j^z,
  &
  \sp{S_\mu,S_\nu}_-
  &= \delta_{\mu,-\nu} 2 \mu S_z.
  \label{eq:comm_mu_nu}
\end{align}
These relations can be used to inductively determine commutation
relations between higher-weight spin operators.  In particular, we can
find that
\begin{align}
  S_z S_\mu^m
  = \p{\mu S_\mu + S_\mu S_z} S_\mu^{m-1}
  = \mu S_\mu^m + S_\mu S_z S_\mu^{m-1}
  = 2 \mu S_\mu^m + S_\mu S_z S_\mu^{m-2}
  = S_\mu^m \p{\mu m + S_z},
  \label{eq:push_z_mu}
\end{align}
and in turn
\begin{align}
  S_z S_\mu^m S_\nu^n
  = S_\mu^m \p{\mu m + S_z} S_\nu^n
  = S_\mu^m S_\nu^n \p{\mu m + \nu n + S_z}.
  \label{eq:push_z_mu_nu}
\end{align}
Repeating this process $\ell$ times tells us
\begin{align}
  S_z^\ell S_\mu^m S_\nu^n
  = S_\mu^m S_\nu^n \p{\mu m + \nu n + S_z}^\ell,
\end{align}
which implies
\begin{align}
  \sp{S_z^\ell, S_\mu^m S_\nu^n}_-
  = S_\mu^m S_\nu^n \p{\sp{\mu m + \nu n + S_z}^\ell - S_z^\ell}
  = S_\mu^m S_\nu^n  \sum_{k=0}^{\ell-1} {\ell \choose k}
  \p{\mu m + \nu n}^{\ell-k} S_z^k.
  \label{eq:comm_z_mu_nu_pow}
\end{align}
Similarly finding commutation relations between high-weight transverse
spin operators, i.e.~$S_\mu^m$ and $S_\nu^n$, turns out to be
considerably more difficult than the cases we have worked out thus
far.  We therefore save this case for Appendix
\ref{sec:comm_transverse}.

\newpage
\section{Commutators between high-weight transverse spin operators}
\label{sec:comm_transverse}

@@

$\bmu\equiv-\mu$ for brevity)

@@

  &= \p{2\mu S_z + S_\bmu S_\mu} S_\bmu^{n-1} \\
  &= 2\mu S_z S_\bmu^{n-1} + S_\bmu S_\mu S_\bmu^{n-1} \\
  &= 2\mu S_z S_\bmu^{n-1}
  + S_\bmu \sp{2\mu S_z S_\bmu^{n-2}
    + S_\bmu S_\mu S_\bmu^{n-2}} \\
  &= 2\mu \p{S_z S_\bmu^{n-1} + S_\bmu S_z S_\bmu ^{n-2}}
  + S_\bmu^2 S_\mu S_\bmu^{n-2} \\
  &= 2\mu \p{S_z S_\bmu^{n-1} + S_\bmu S_z S_\bmu ^{n-2}
    + S_\bmu^2 S_z S_\bmu ^{n-3}} + S_\bmu^3 S_\mu S_\bmu^{n-3} \\
  &= 2\mu \sum_{k=0}^{n-1} S_\bmu^k S_z S_\bmu^{n-k-1}
  + S_\bmu^n S_\mu.
  \label{eq:comm_mu_nu_single}

@@

While \eqref{eq:comm_mu_nu_single} gives us the commutator

@@

  &= S_\mu^{m-1} \sp{n S_\bmu^{n-1}\p{1 - n + 2\mu S_z}
    + S_\bmu^n S_\mu} \\
  &= n \sp{S_\mu^{m-1} S_\bmu^{n-1} \p{1 - n + 2\mu S_z}}
  + S_\mu^{m-1} S_\bmu^n S_\mu \\

>>>>

Previous work has shown how to realize various spin-squeezing
Hamiltonians with ultracold atomic systems.  While it is
straightforward to simulate the dynamics generated by these
Hamiltonians in the context of a closed quantum system, here we try to
develop an understanding of their the squeezing behavior in the
presence of decoherence.  In particular, we will be considering
dynamics induced by the collective one-axis twisting (OAT), two-axis
twisting (TAT), and transverse field (TVF) Hamiltonians of the form
\begin{align}
  H_{\t{OAT}} = \chi S_z^2,
  &&
  H_{\t{TAT}} = \chi \p{S_z^2 - S_y^2},
  &&
  H_{\t{TVF}} = \chi S_z^2 - \Omega S_x,
\end{align}
for spins which decay in an uncorrelated manner at a rate $\gamma$.
The evolution of any Heisenburg operator $\O$ with a Hamiltonian $H$
is given by
\begin{align}
  \f{d}{dt} \O
  = i\sp{H,\O}_- + \gamma\D_-\p{\O},
  &&
  \D_-\p{\O} \equiv \sum_j\p{\sigma_j^+\O\sigma_j^-
    - \f12\sp{\sigma_j^+\sigma_j^-,\O}_+},
  \label{eq:EOM}
\end{align}
where $\sp{A,B}_\pm\equiv AB\pm BA$ and $j$ indexes an individual
spin.  Our measure of squeezing is
\begin{align}
  \xi^2 \equiv \f{N}{\abs{\bk{\v S}}^2}
  \min_{\uv n\perp\bk{\v S}} \bk{\p{\v S\c\uv n}^2},
  \label{eq:squeezing}
\end{align}
where $N$ is the total number of spins, $\v S$ is the collective spin
vector, and the minimization is performed over all unit vectors
$\uv n$ in the plane orthogonal to the mean spin vector $\bk{\v S}$.
This squeezing parameter is entirely determined by components
$\bk{S_\alpha}$ of the mean spin vector in addition to collective
spin-spin correlators of the form $\bk{S_\alpha S_\beta}$; our task is
therefore to compute these expectation values in all scenarios of
interest.

@@

The case of the one-axis twisting is simplest to consider, as we can
write the Hamiltonian in the form
\begin{align}
  H_{\t{OAT}} = \f14 \chi \sum_{j,k} \sigma_j^z \sigma_k^z
  = \f12 \chi \sum_{j<k} \sigma_j^z \sigma_k^z + \f14 N \chi,
\end{align}
which admits an exact solution previously worked out by Foss-Feig et
al.~in ref.~\cite{foss-feig2013nonequilibrium}.  Adapting exact
expectation values with respect to the initial state $\ket{X}$ which
has all spins pointing in $x$, in terms of $\mu,\nu\in\set{+1,-1}$ we
find that
\begin{align}
  \bk{S_+}
  &= \sum_j\bk{\sigma_j^+}
  = \f{N}{2} e^{-\gamma t/2} \Phi\p{\chi,t}^{N-1}, \label{eq:S+_OAT} \\
  \bk{S_\mu S_z}
  &= \f12\sum_j\bk{\sigma_j^\mu\sigma_j^z}
  + \f12\sum_{j\ne k} \bk{\sigma_j^\mu \sigma_k^z} \\
  &= -\f{\mu}{2}\bk{S_\mu} + \f14 N \p{N-1} e^{-\gamma t/2}
  \Psi\p{\mu\chi,t} \Phi\p{\chi,t}^{N-2}, \\
  \bk{S_\mu S_\nu}
  &= \sum_j \bk{\sigma_j^\mu \sigma_j^\nu}
  + \sum_{j\ne k} \bk{\sigma_j^\mu \sigma_k^\nu} \\
  &= \delta_{\mu,-\nu} \p{\f{N}{2} + \mu \bk{S_z}}
  + \f14 N \p{N-1} e^{-\gamma t}
  \Phi\p{\sp{\mu+\nu}\chi,t}^{N-2}, \label{eq:SS+-_OAT}
\end{align}
where
\begin{align}
  \Phi\p{\zeta,t} \equiv e^{-\gamma t/2}
  \sp{\cos\p{s_\zeta t}+\f12\gamma t~\t{sinc}\p{s_\zeta t}},
  &&
  \Psi\p{\zeta,t} \equiv e^{-\gamma t/2}
  \p{is_\zeta-\gamma/2}t~ \t{sinc}\p{s_\zeta t},
\end{align}
and $s_\zeta \equiv \zeta + i2\gamma$.  The last ingredients we need
to compute spin squeezing at any time are $\bk{S_z}$ and $\bk{S_z^2}$.
As these operators commute with the OAT Hamiltonian, their evolution
is governed entirely by decoherence, such that
\begin{align}
  \f{d}{dt}\p{S_z^n}
  = \gamma \D_-\p{S_z^n}
  = \gamma \p{S + S_z} \sum_{k=0}^{n-1}
  \p{-1}^{n-k} { n \choose k } S_z^k,
\end{align}
where $S\equiv N/2$ and $\D_-\p{S_z^n}$ is worked out in Appendix
\ref{sec:decoherence}.  In particular,
\begin{align}
  \f{d}{dt} S_z = -\gamma \p{S + S_z},
  &&
  \f{d}{dt}\p{S_z^2} = \gamma \p{S + S_z} \p{1 - 2S_z}.
\end{align}
The initial conditions $\bk{S_z}_{t=0}=0$ and $\bk{S_z^2}_{t=0}=S/2$
then imply
\begin{align}
  \bk{S_z} = S\p{e^{-\gamma t}-1},
  &&
  \bk{S_z^2} = S \p{1 - \f12 e^{-\gamma t}} e^{-\gamma t} + \bk{S_z}^2.
  \label{eq:Sz_OAT}
\end{align}
The expectation values in \eqref{eq:S+_OAT}-\eqref{eq:SS+-_OAT} and
\eqref{eq:Sz_OAT} are sufficient to compute the squeezing parameter
$\xi^2$ in \eqref{eq:squeezing} at any time throughout evolution of
the initial state $\ket{X}$ under $H_{\t{OAT}}$.

@@

The equations of motion for $S_\mu$, $S_\mu S_z$, and $S_\mu S_\nu$
(i.e.~where $\mu,\nu\in\set{+1,-1}$) under the OAT Hamiltonian
$H_{\t{OAT}}=\chi S_z^2$ are determined by
\begin{align}
  \f{d}{dt} \p{S_+^\ell S_-^m S_z^n}
  = i\chi \sp{S_z^2, S_+^\ell S_-^m S_z^n}_-
  + \gamma \D_-\p{S_+^\ell S_-^m S_z^n},
\end{align}
where we can use identity \eqref{eq:comm_z_mu} to simplify
\begin{align}
  \sp{S_z^2, S_+^\ell S_-^m S_z^n}_-
  = \sp{S_z^2, S_+^\ell S_-^m}_- S_z^n
  = S_+^\ell S_-^m \p{\sp{\ell-m}^2 + 2\sp{\ell-m} S_z} S_z^n,
  \label{eq:S_z^2_comm}
\end{align}
and the effect of decoherence through $\D_-\p{S_+^\ell S_-^m S_z^n}$ is
worked out in Appendix \ref{sec:decoherence}.  In order to compute
squeezing of the OAT Hamiltonian in the presence of single-spin decay,
we therefore need to solve the systems of equations defined by
\begin{align}
  \f{d}{dt} \p{S_\mu S_z^n}
  = i\chi S_\mu \p{1 + 2\mu S_z} S_z^n
  + \gamma S_\mu \sp{\p{S - \delta_{\mu,-1} + S_z} \sum_{k=0}^{n-1}
    \p{-1}^{n-k} { n \choose k } S_z^k - \f12 S_z^n}
\end{align}
and
\begin{multline}
  \f{d}{dt} \p{S_\mu S_\nu S_z^n}
  = i\chi S_\mu S_\nu \p{\sp{\mu+\nu}^2 + 2 \sp{\mu+\nu} S_z} S_z^n \\
  + \gamma S_\mu S_\nu
  \sp{\p{S - \delta_{\mu,-1} - \delta_{\nu,-1} + S_z} \sum_{k=0}^{n-1}
    \p{-1}^{n-k} { n \choose k } S_z^k - S_z^n},
\end{multline}
for $\mu\ge\nu$.  The initial conditions for these systems of equations can be determined from the expansions
\begin{align}
  S_\mu = \sum_{m=-S}^S \sqrt{\p{S-\mu m}\p{S+\mu m+1}} \op{m+\mu}{m},
  &&
  S_z^n = \sum_{m=-S}^S m^n \op{m},
\end{align}
and the initial state
\begin{align}
  \ket{X} = 2^{-N/2} \sum_{m=-S}^S { N \choose S+m }^{1/2} \ket{m},
\end{align}
from which we can determine
\begin{align}
  \bk{S_\mu S_z^n}_{t=0}
  &= 2^{-N} \sum_{m=-S}^S m^n
  \sp{{ N \choose S+m } { N \choose S+m+\mu }
    \p{S-\mu m}\p{S+\mu m+1}}^{1/2} \\
  &= 2^{-N} \sum_{m=-S}^S m^n { N \choose S+m } \abs{m - \mu S},
\end{align}
and
\begin{multline}
  \bk{S_\mu S_\nu S_z^n}_{t=0}
  = 2^{-N} \sum_{m=-S}^S m^n
  \sp{{ N \choose S+m } { N \choose S+m+\mu+\nu }}^{1/2} \\
  \times \sp{\p{S-\nu\sp{m+\nu}}\p{S+\mu\sp{m+\nu}+1}
    \p{S-\nu m}\p{S+\nu m+1}}^{1/2}.
\end{multline}

% \section{One-axis twisting with a transverse field}

% We now consider again squeezing the initial state $\ket{X}$, but this
% time with the transverse field Hamiltonian
% \begin{align}
%   H_{\t{TVF}}
%   \equiv \chi S_z^2- \Omega S_x
%   = \chi S_z^2- \f{\Omega}{2} \p{S_+ + S_-}.
% \end{align}
% To determine the coherent evolution of the operator
% $S_+^\ell S_-^m S_z^n$, we now need to compute
% \begin{align}
%   \sp{H_{\t{TVF}}, S_+^\ell S_-^m S_z^n}_-
%   = \chi \sp{S_z^2, S_+^\ell S_-^m S_z^n}_-
%   - \f{\Omega}{2} \sp{S_+, S_+^\ell S_-^m S_z^n}_-
%   - \f{\Omega}{2} \sp{S_-, S_+^\ell S_-^m S_z^n}_-.
% \end{align}
% The first commutator above is identical to that for OAT, and is given
% in \eqref{eq:S_z^2_comm}.  For the second commutator, we expand
% \begin{align}
%   \sp{S_+, S_+^\ell S_-^m S_z^n}_-
%   = S_+^{\ell+1} S_-^m S_z^n - S_+^\ell S_-^m S_z^n S_+,
% \end{align}
% where
% \begin{align}
%   S_-^m S_z^n S_+
%   = S_-^m S_+ \p{1 + S_z}^n
%   = \sp{S_+ S_-^m - m S_-^{m-1} \p{1 - m + 2 S_z}}
%   \sum_{k=0}^n { n \choose k } S_z^k,
% \end{align}
% so
% \begin{align}
%   \sp{S_+, S_+^\ell S_-^m S_z^n}_-
%   = S_+^{\ell+1} S_-^m S_z^n
%   - S_+^\ell \sp{S_+ S_-^m - m S_-^{m-1} \p{1 - m + 2 S_z}}
%   \sum_{k=0}^n { n \choose k } S_z^k.
% \end{align}
% Similarly,
% \begin{align}
%   \sp{S_-, S_+^\ell S_-^m S_z^n}_-
%   = S_- S_+^\ell S_-^m S_z^n - S_+^\ell S_-^m S_z^n S_-,
% \end{align}
% where
% \begin{align}
%   S_- S_+^\ell S_-^m
%   = S_+^\ell S_-^{m+1} + \ell S_+^{\ell-1}\p{1-\ell-2S_z} S_-^m
%   = S_+^\ell S_-^{m+1} + \ell S_+^{\ell-1} S_-^m \p{1-m-\ell-2S_z},
% \end{align}
%   and
% \begin{align}
%   S_z^n S_- = S_- \p{1 - S_z}^n
%   = S_- \sum_{k=0}^n \p{-1}^k { n \choose k } S_z^k,
% \end{align}
% so
% \begin{align}
%   \sp{S_-, S_+^\ell S_-^m S_z^n}_-
%   = S_+^\ell S_-^{m+1} S_z^n
%   + \ell S_+^{\ell-1} S_-^m \p{1-m-\ell-2S_z} S_z^n
%   - S_+^\ell S_-^{m+1} \sum_{k=0}^n \p{-1}^k { n \choose k } S_z^k.
% \end{align}

\section{Evolution under general Hamiltonians with decoherence}

All collective spin Hamiltonians can be written in the form
\begin{align}
  H = \sum_{p,q,r} h_{pqr} S_+^p S_-^q S_z^r,
\end{align}
for which the coherent evolution of a general operator
$S_+^\ell S_-^m S_z^n$ is given by
\begin{align}
  \f{d}{dt} \p{S_+^\ell S_-^m S_z^n}
  = i\sum_{p,q,r} h_{pqr} \sp{S_+^p S_-^q S_z^r, S_+^\ell S_-^m S_z^n}_-
  + \D_-\p{S_+^\ell S_-^m S_z^n}.
\end{align}

@@

This text makes ubiquitous use of various spin operator identities; we
collect and derive particularly useful identities here for reference.
The elementary commutation relations between single-weight spin
operators are, with $\bmu\equiv-\mu\in\set{+1,-1}$ for brevity,
\begin{align}
  \f12 \sp{\sigma_j^z,\sigma_k^\mu}_-
  &= \delta_{jk} \mu \sigma_j^\mu,
  &
  \sp{S_z,\sigma_j^\mu}_-
  &= \f12 \sp{\sigma_j^z,S_\mu}_- = \mu \sigma_j^\mu,
  &
  \sp{S_z,S_\mu}_-
  &= \mu S_\mu,
  \label{eq:comm_z_mu_base} \\
  \sp{\sigma_j^\mu,\sigma_k^\bmu}_-
  &= \delta_{jk} \mu \sigma_j^z,
  &
  \sp{S_\mu,\sigma_j^\bmu}_-
  &= \sp{\sigma_j^\mu,S_\bmu}_- = \mu \sigma_j^z,
  &
  \sp{S_\mu,S_\bmu}_-
  &= 2 \mu S_z.
  \label{eq:comm_mu_base}
\end{align}
These relations can be used to inductively determine commutation
relations between higher-weight spin operators.  In particular, we can
find that
\begin{align}
  S_z S_\mu^m
  = \p{\mu S_\mu + S_\mu S_z} S_\mu^{m-1}
  = \mu S_\mu^m + S_\mu S_z S_\mu^{m-1}
  = 2 \mu S_\mu^m + S_\mu S_z S_\mu^{m-2}
  = S_\mu^m \p{\mu m + S_z},
  \label{eq:push_z_mu_single}
\end{align}
and in turn
\begin{align}
  S_z S_\mu^m S_\bmu^n
  = S_\mu^m \p{\mu m + S_z} S_\bmu^n
  = S_\mu^m S_\bmu^n \p{\mu\sp{m - n} + S_z}.
\end{align}
Repeating this process $\ell$ times tells us
\begin{align}
  S_z^\ell S_\mu^m S_\bmu^n
  = S_\mu^m S_\bmu^n \p{\mu\sp{m - n} + S_z}^\ell,
  \label{eq:push_z_mu}
\end{align}
which implies
\begin{align}
  \sp{S_z^\ell, S_\mu^m S_\bmu^n}_-
  = S_\mu^m S_\bmu^n \sp{\mu^\ell \p{\mu\sp{m - n} + S_z}^\ell - S_z^\ell}
  = S_\mu^m S_\bmu^n  \sum_{k=0}^{\ell-1} {\ell \choose k}
  \mu^{\ell-k} \p{m-n}^{\ell-k} S_z^k.
  \label{eq:comm_z_mu}
\end{align}
Similarly finding commutation relations between high-weight transverse
spin operators, i.e.~$S_\mu^m$ and $S_\bmu^n$, turns out to be
considerably more difficult than the cases we have worked out thus
far.  We therefore save this case for Appendix
\ref{sec:comm_transverse}.

\section{Decoherence of Heisenberg spin operators}
\label{sec:decoherence}

The decoherence operator for uncorrelated incoherent decay ($\mu=-1$)
or excitation ($\mu=1$) of individual spins is, with $\bmu\equiv-\mu$,
\begin{align}
  \D_\mu\p{\O}
  \equiv \sum_j\p{\sigma_j^\bmu \O \sigma_j^\mu
    - \f12\sp{\sigma_j^\bmu \sigma_j^\mu,\O}_+},
\end{align}
where
\begin{align}
  \sum_j \f12\sp{\sigma_j^\bmu \sigma_j^\mu,\O}_+
  = \f12\sp{S-\mu S_z,\O}_+
  = S\O - \f{\mu}{2}\sp{S_z,\O}_+
  = \O\p{S-\mu S_z} - \f{\mu}{2}\sp{S_z,\O}_-,
\end{align}
so
\begin{align}
  \D_\mu\p{\O}
  = \sum_j \sigma_j^\bmu \O \sigma_j^\mu
  - \O\p{S - \mu S_z} + \f{\mu}{2} \sp{S_z,\O}_-.
\end{align}
We are interested in the effect of this decoherence on the dynamics of
collective spin operators of the form $S_\bmu^\ell S_\mu^m S_z^n$.  We
thus expand
\begin{align}
  \D_\mu\p{S_\bmu^\ell S_\mu^m S_z^n}
  = \sum_j \sigma_j^\bmu S_\bmu^\ell S_\mu^m S_z^n \sigma_j^\mu
  - S_\bmu^\ell S_\mu^m S_z^n \p{S - \mu S_z}
  + \f{\mu}{2} \sp{S_z, S_\bmu^\ell S_\mu^m S_z^n}_-.
  \label{eq:D_z+-}
\end{align}
The commutator in \eqref{eq:D_z+-} is straightforward to simplify
using \eqref{eq:comm_z_mu}:
\begin{align}
  \mu \sp{S_z, S_\bmu^\ell S_\mu^m S_z^n}_-
  = \mu \sp{S_z, S_\bmu^\ell S_\mu^m}_- S_z^n
  = \mu\bmu \p{\ell-m} S_\bmu^\ell S_\mu^m S_z^n
  = - \p{\ell-m} S_\bmu^\ell S_\mu^m S_z^n,
\end{align}
while the first term in \eqref{eq:D_z+-} can be written in terms of
collective spin operators if we use
\eqref{eq:comm_z_mu_base}-\eqref{eq:comm_mu_base} to push the jump
operators together:
\begin{align}
  \sum_j \sigma_j^\bmu S_\bmu^\ell S_\mu^m S_z^n \sigma_j^\mu
  = \sum_j S_\bmu^\ell \sigma_j^\bmu \sigma_j^\mu S_\mu^m \p{\mu+S_z}^n
  = S_\bmu^\ell \p{S - \mu S_z} S_\mu^m
  \sum_{k=0}^n \mu^{n-k} { n \choose k } S_z^k.
\end{align}
Finally, we need to push the middle $S_z$ to the right, which we can
do using \eqref{eq:push_z_mu_single}:
\begin{align}
  \sum_j \sigma_j^\bmu S_\bmu^\ell S_\mu^m S_z^n \sigma_j^\mu
  = S_\bmu^\ell S_\mu^m \p{S - m - \mu S_z}
  \sum_{k=0}^n \mu^{n-k} { n \choose k } S_z^k.
\end{align}
Altogether, we have that
\begin{align}
  \D_\mu\p{S_\bmu^\ell S_\mu^m S_z^n}
  &= S_\bmu^\ell S_\mu^m \sp{\p{S - m - \mu S_z} \sum_{k=0}^n
    \mu^{n-k} { n \choose k } S_z^k
    - \p{S - \mu S_z + \f12\sp{\ell-m}} S_z^n} \\
  &= S_\bmu^\ell S_\mu^m \sp{\p{S - m - \mu S_z} \sum_{k=0}^{n-1}
    \mu^{n-k} { n \choose k } S_z^k - \f12 \p{\ell+m} S_z^n}.
\end{align}

\section{Commutators between high-weight transverse spin operators}
\label{sec:comm_transverse}

@@

$\bmu\equiv-\mu\in\set{+1,-1}$)

@@

  &= S_\bmu S_\mu S_\bmu^{n-1} + 2\mu S_z S_\bmu^{n-1} \\
  &= S_\bmu \sp{S_\bmu S_\mu S_\bmu^{n-2} + 2\mu S_z S_\bmu^{n-2}}
  + 2\mu S_z S_\bmu^{n-1} \\
  &= S_\bmu^2 S_\mu S_\bmu^{n-2}
  + 2\mu \p{S_\bmu S_z S_\bmu ^{n-2} + S_z S_\bmu^{n-1}} \\
  &= S_\bmu^n S_\mu + 2\mu \sum_{k=0}^{n-1} S_\bmu^k S_z S_\bmu^{n-k-1}.
  \label{eq:comm_mu_single}

@@

While \eqref{eq:comm_mu_single} gives us the commutator

@@

  &= S_\mu^{m-1} S_\bmu^n S_\mu
  + n \sp{S_\mu^{m-1} S_\bmu^{n-1} \p{1 - n + 2\mu S_z}} \\
  &= S_\bmu^n S_\mu^m
  + n \sum_{k=0}^{m-1} S_\mu^{m-k-1} S_\bmu^{n-1}
  \p{1 - n + 2\mu S_z} S_\mu^k,
